Leadership across cultures and contexts: an exploratory analysis

### 2020-05-15 leadership_across_cultures_contexts.Rmd
Pulled out most WEIRD discussions from main text and put in SI.
Minor change.

<!-- WEIRD PEOPLE PROBLEMS -->

There is now broad consensus within the behavioral sciences on the importance of data from the full range of human cultural diversity. Systemic biases in much of social science are well known, and might be due, in part, to the lack of data from populations other than American undergraduates [@henrich_weirdest_2010, *The weirdest people in the world?*]. Since the seminal WEIRD people paper, many social scientists have taken up the challenge to diversify their samples and have sought out "non-WEIRD" participants and databases. Dichotomizing populations and their psychologies and social dynamics as WEIRD or non-WEIRD, however, is a mistake.^[A Google Scholar search for "non-WEIRD" produced over 1000 results.] As anthropology has conclusively demonstrated, there is enormous diversity across societies that does not remotely resemble a dichotomy.

@@

<!-- CLOSING -->
We first briefly review theoretical, ethnographic, and empirical perspectives on the qualities of leaders, the functions they serve, and the costs and benefits of leadership across human societies, including WEIRD and non-WEIRD populations. We then outline the aims and methods of the current cross-cultural study -- which focuses on exploring patterns among non-WEIRD populations -- before presenting results and interpretations. Through novel exploratory analyses encompassing a broad range of cultural diversity and social contexts we aim to advance leadership studies across disciplines and provide researchers with a foundation for more detailed approaches to human leadership across cultures and contexts.

@@

<!-- MORE UNIVERSAL QUALITIES FROM WEIRD AND NON WEIRD -->
Across cultures elders, respected individuals, and highly intelligent or knowledgeable individuals tend to have greater influence over community decisions including among the most egalitarian hunter-gatherers, stratified chiefdoms, and contemporary state-level organizations [@silverman_how_1978; @van_vugt_cognitive_2007; @tooby_cognitive_2006; @garfield_review_2019; @bass_handbook_1990]. Drawing on cross-national data through the GLOBE project from over 15,000 middle managers in businesses across 60 countries, @den_hartog_culture_1999 conclude that elements of charismatic and transformational leadership -- leadership styles when leaders express an inclusive vision of future progress, lead through inspiration, and develop initiatives in collaboration -- are universally endorsed across nations and are qualities perceived to be associated with effective leadership. In particular, inducing motivation, having foresight, being encouraging and communicative, trustworthiness, and building self-confidence among followers are suggested to be universal attributes of outstanding leadership in managerial contexts.

@@

<!-- INTELLIGENCE -->
Many theoretical models and empirical findings have implicated high intelligence as an important quality positively associated with effective leadership [e.g., @johnson_leadership_2015; @judge_intelligence_2004; @neel_being_1980; @antonakis_does_2009], including across diverse societies [@garfield_evolutionary_2019; @garfield_investigating_2019]. These results suggests high intelligence and specialized knowledge are potentially universal traits of political and community leaders. Within managerial studies scholars often suggest the cognitive demands of leadership escalate with increasing organizational complexity [@zaccaro_organizational_2002]. More complex organizations are thought to require leaders equipped to process complex information, understand distinct departmentalized bodies, and resolve conflicting points of view among diverse individuals [@bass_handbook_1990; @schneider_organizations_2006; @antonakis_does_2009]. Some recent evidence, however, suggests a curvilinear effect in that exceptionally intelligence individuals at the high end of the distribution may not be the most effective leaders in contemporary business settings [@antonakis_can_2017].

<!--  CONCLUSION PARAGRAPH/STATEMENT. -->
Much of the theoretical leadership literature across disciplines suggests relative uniformity in many qualities of community leaders. Intelligence, charisma, effective decision-making capabilities, high social status and strong social networks are likely to be broadly representative qualities of political and community leaders. In more socially stratified, politically complex, and in increasingly WEIRD contexts, heredity, ascribed social statuses and affiliations are more likely to be critical qualities in achieving and/or maintaining leadership positions.

@@

<!-- NEED CONCLUSTION STATEMENT/PARAGRAPH.  ?????-->

@@

>>>>

<!-- WEIRD PEOPLE PROBLEMS -->
There is now broad consensus within the behavioral sciences on the importance of data from the full range of human cultural diversity. Systemic biases in much of social science are well known, and might be due, in part, to the lack of data from populations other than American undergraduates [@henrich_weirdest_2010, *The weirdest people in the world?*]. Since the seminal WEIRD people paper, many social scientists have taken up the challenge to diversify their samples and have sought out "non-WEIRD" participants and databases. Dichotomizing populations and their psychologies and social dynamics as WEIRD or non-WEIRD, however, is a mistake.^[A Google Scholar search for "non-WEIRD" produced over 1000 results.] As anthropology has conclusively demonstrated, there is enormous diversity across societies that does not remotely resemble a dichotomy.

@@

<!-- CLOSING -->
We first briefly review theoretical, ethnographic, and empirical perspectives on the qualities of leaders, the functions they serve, and the costs and benefits of leadership across nonindustrial non-WEIRD human societies (we include detiled review of related literature on WEIRD populations in the SI). We then outline the aims and methods of the current cross-cultural study -- which focuses on exploring patterns among non-WEIRD populations -- before presenting results and interpretations. Through novel exploratory analyses encompassing a broad range of cultural diversity and social contexts we aim to advance leadership studies across disciplines and provide researchers with a foundation for more detailed approaches to human leadership across cultures and contexts.

@@

<!-- MORE UNIVERSAL QUALITIES FROM WEIRD AND NON WEIRD -->
Across cultures elders, respected individuals, and highly intelligent or knowledgeable individuals tend to have greater influence over community decisions including among the most egalitarian hunter-gatherers, stratified chiefdoms, and contemporary state-level organizations [@silverman_how_1978; @van_vugt_cognitive_2007; @tooby_cognitive_2006; @garfield_review_2019; @bass_handbook_1990].

@@

<!-- INTELLIGENCE -->
Many theoretical models and empirical findings have implicated high intelligence as an important quality positively associated with effective leadership [e.g., @johnson_leadership_2015; @judge_intelligence_2004; @neel_being_1980; @antonakis_does_2009], including across diverse societies [@garfield_evolutionary_2019; @garfield_investigating_2019]. These results suggests high intelligence and specialized knowledge are potentially universal traits of political and community leaders.

<!--  CONCLUSION PARAGRAPH/STATEMENT. -->
Much of the theoretical leadership literature across disciplines suggests relative uniformity in many qualities of community leaders. Intelligence, charisma, effective decision-making capabilities, high social status and strong social networks are likely to be broadly representative qualities of political and community leaders.

@@

@@

## WEIRD leadership dimensions

### WEIRD leadership functions

### WEIRD leadership qualities

Drawing on cross-national data through the GLOBE project from over 15,000 middle managers in businesses across 60 countries, @den_hartog_culture_1999 conclude that elements of charismatic and transformational leadership -- leadership styles when leaders express an inclusive vision of future progress, lead through inspiration, and develop initiatives in collaboration -- are universally endorsed across nations and are qualities perceived to be associated with effective leadership. In particular, inducing motivation, having foresight, being encouraging and communicative, trustworthiness, and building self-confidence among followers are suggested to be universal attributes of outstanding leadership in managerial contexts.

Within managerial studies scholars often suggest the cognitive demands of leadership escalate with increasing organizational complexity [@zaccaro_organizational_2002]. More complex organizations are thought to require leaders equipped to process complex information, understand distinct departmentalized bodies, and resolve conflicting points of view among diverse individuals [@bass_handbook_1990; @schneider_organizations_2006; @antonakis_does_2009]. Some recent evidence, however, suggests a curvilinear effect in that exceptionally intelligence individuals at the high end of the distribution may not be the most effective leaders in contemporary business settings [@antonakis_can_2017].

In more socially stratified, politically complex, and in increasingly WEIRD contexts, heredity, ascribed social statuses and affiliations are more likely to be critical qualities in achieving and/or maintaining leadership positions.

### WEIRD costs and benefits
